\normalfont\spacedallcaps{Hazardous asteroids forecast via Markov random fields}

### 2021-11-29 Manuscript/manuscript.tex
PUB

\section*{Abstract} The machine learning algorithms provide a promising approach to classify and predict natural or social phenomena. Differently to the oldest theoretical approach in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynmanlectures}), the machine learning algorithms start from very general assumptions/expressions and then their parameters, are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}. The first approach allows obtaining interpretable predictions, the second one good forecast with a reduced effort. However, the toll that has to be paid in this case is the low interpretability \cite{russell2010artificial,murphy2012machine}. Since they provide the list of conditional independences/dependencies, the graphical methods represent a good compromise between the need for interpretability and the forecasts obtained without developing a general theory. In order to prove the validity of this statement, I considered a dataset for which the theory that interconnects its feature was known: in particular, I chose the asteroids hazardousness as provided by CNEOS \cite{cneos+nasa} and published on Kaggle \cite{kaggle_dataset}. The outcomes of the probabilistic methods for predicting the asteroid hazardousness were compared with the ones provided by the theory and with the ones of Random Forest (RF), Support Vector Machines (SVM), and Logistic Regression and Quadric Discriminant analysis (QDA). The results show that the forecast performances of probabilistic methods are better than QDA and almost equal to the logistic regression but lower to RF and SVM. However, since the list of conditional dependencies correctly reflects the laws of celestial mechanics \cite{murray1999solar}, it can be said that in this case, the probabilistic methods provide an interpretable and correct explanation of their mechanism. Therefore, contrary to other machine learning algorithms, such methods can be fully validated scientifically.

@@

\chapter{Introduction} The description and the forecast of physical \footnote{Where here physical should be intended as mesuarable} phenomena can be attacked with two different approaches: starting from a restricted set of principles/axioms, one can formulate theoretical models that provide equations that describe such phenomena. Such approach will be called here \textit{ab-initio}. On the other side, one can start from data and fit them into some general models (with fixed or not fixed number of parameters). These methods will be called here the machine learning (ML) methods. In the ab-initio case, one can fully explain the laws obtained and an overall idea of why nature works in this way. The toll that has to be paid in this case is that, since the phenomena are much more complicated than the starting principles, a consistent part of them may be excluded by the assumption made at the beginning.
The drawback of this approach is that the underlying mechanism by which a forecast is preferred to another one may not be interpretable. Furthermore, the calculation of the solution given the dynamical equation can be computationally expensive\footnote{This is the case, for instance, of the Hartree-Fock method or Density Functional Theory: such methods rewrite, with some approximations, the Schroedinger equation into a much more computationally affordable way. However, also, in this case, the solution for solids can be costly \cite{martin_2004}}. On the other side, the ML methods, since they make much more general assumptions and work directly on data, are not limited to a particular class of phenomena. Therefore they do not require the development of a specific theory as the \textit{ab-initio} one: the same model can be applied to exoplanets habitability as well to credit risk evaluation. Such dichotomy between these two methods can be explained with an example taken from astronomy: at the beginning of XX century, astronomers knew that Mercury follows an elliptical orbit that, instead of being fixed, rotates around the Sun as shown in Fig. \ref{Perihelion_precession2} Such phenomenon is called \textit{Precession of the perihelion of Mercury}. The crucial point is that there was no way to explain from the Newton gravity theory such phenomenon. At this point, scholars had two paths: describe its motion with an empirical law (perhaps with an empirical modification of Newton's gravity law) or reformulate the Newton gravity from scratch. The second approach was the one followed by A.Einstein with his general relativity theory (further details are given in the A.Zee textbook \cite{zee2013einstein}). The graphical methods represent a good compromise between these two approaches: they can be used for every physical phenomenon, and, since they provide conditional dependencies/independences, they also give an elegant and intuitive representation of their mechanisms. Therefore, if the theory of the involved process is known, one can use this class of algorithms and compare its finding with the theory. In this way, one can evaluate the quality of the forecasts and the model provided by the graphical methods. The present work aims to perform such analysis for a dataset that contain the features and the hazardousness \footnote{For Earth} of asteroids. For this dataset, indeed, the relationships between the various features are known. The present work is organized as follows: first, a brief introduction about the basic theoretical concepts of the statistical methods here used will be provided, then the dataset will be described, and finally, the principal results here obtained with graphical methods will be presented as compared with other types of machine learning algorithms. Furthermore, a recap of the celestial mechanics will be provided in Appendix A. This theory explains the connections and the meaning of the different features contained in the dataset.

@@

@@

\section{Information theory}
Given an ensemble of random variables, we can quantify the amount of information that one variable contains: such quantity is called mutual information and is a key concept within information theory. This approach, which was implemented by Claude Shannon decades before the probabilistic modelling, represents a complementary way by which one can attack the problem of conditional dependence between random variables. Here we would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allow defining the concept of mutual information properly. The founding concept of information theory is entropy. This quantity expresses the uncertainty of a random variable. Given a random variable $X$ with alphabet (the accessible states) $\cite{cover2006elements}$ and probability mass function $p(x)=Pr\left\lbrace X=x \rbrace\; x \in \chi$ we define the entropy of $X$ as $H(X)=-\sum_{x \in X} p(x)log p(x)$ where the logarithm has to be considered with basis 2. In analogous way the joint entropy of two random variables $(X,Y)$ with a joint distribution p(x,y) is defined as \cite{cover2006elements}:
\begin{equation}
H(X,Y)=-\sum_{x\in \chi}\sum_{y\in \mathcal{Y}}p(x,y)\log p(x,y)
\end{equation}
Furthermore, we can define also the conditional entropy as \cite{cover2006elements}:
\begin{equation}
\begin{split}
H(X|Y)=& \sum_{x\in \mathcal{X} }p(x)H(Y|X=x)\\
=& -\sum_{x\in \mathcal{X}}p(x)\sum_{y\in \mathcal{Y}}p(x,y)\log p(y|x) \\
=& -E\log p(Y|X)
\end{split}
\end{equation}
The joint entropy and the conditional entropy are related by the chain rule \cite{cover2006elements}:
\begin{equation}
H(X,Y)=H(X)+H(Y|X)
\end{equation}
Such rule can be extended to to the following from \cite{cover2006elements}:
\begin{equation}
H(X,Y|Z)=H(X|Z)+H(Y|X,Z)
\end{equation}
Given a distribution q and another distribution p, one can quantify how inefficiently the second one describes the first one using the concept of relative entropy or Kullback-Leibler distance \cite{cover2006elements,mackay2003information}:
\begin{equation}
D(p||q)=\sum p(x)\log\frac{p(x)}{q(x)}
\end{equation}
As stated by the Gibbs inequality \cite{cover2006elements,mackay2003information}:
\begin{equation}
D(p||q)\geq 0
\end{equation}
this quantity can not be negative: the entropy of a random variable associated with another cannot have a degree of uncertainty lower to the quantity it aims to describe. On these bases, we are now ready to introduce the concept of mutual information. This is defined as  \cite{cover2006elements}:
\begin{equation}
\begin{split}
I(X;Y)&=\sum\sump(x,y)\log\dfrac{p(x,y)}{p(x)p(y)}=\\
&=D(p(x,y)||p(x)p(y)) \\
&=H(X)-H(X|Y)=H(Y)-H(Y|X)
\end{split}
\end{equation}
As for the joint distribution also in this case, we have a chain rule  \cite{cover2006elements}:
\begin{equation}
I(X_{1},X_{2},...,X_{n};Y)=\sum^{n}_{i=1}I(X_{i};Y|X_{i-1},X_{i-2},...,X_{1})
\end{equation}
Finally, we would report the data process inequality theorem that connects the information theory with the Markov chain: if we have a Markov chain, $X\rightarrow Y \rightarrow Z$  then $I(X;Y)\geq I(X;Z) $. As for the Gibbs inequality, the underlying idea is that no clever manipulation of the data can improve the inference that can be made on them \cite{cover2006elements,mackay2003information}. Otherwise, we would have a clear violation of the second principle of thermodynamics (see for instance, the Maxwell's demon \cite{feynman2018feynman}).

\section{Algorithm intepretability}

@@

@@

The analysis with the graphical models was started by inspecting the relations between continuous variables in the dataset. Therefore, in this first step, the binary variable \textit{Hazardous} was excluded. For this purpose, among the different methods available, we considered the \textit{Graphical least absolute shrinkage and selection operator} GLASSO as implemented in the \textit{glasso} R package \cite{friedman2008sparse,glasso}. After different tests reported in Fig. \ref{GLASSO_convergence}, we considered as the final result the graph obtained with the value of $\rho$ (the one that penalizes further connections) equal to $0.3$. Such choice is motivated by the fact that the conditional dependences/independences stated by the Celestial Mechanics are correctly reproduced with this parameter. For instance, we see that the diameter of the asteroids is independent of all features related to their motion. Such a result is meaningful since the volume/mass of the asteroids is largely lower than the mass of the earth.  Thus, as explained in Appendix A, there is no way the asteroid orbit can be modified by its mass. In addition, we see that the Close approach Date and the Epoch date are linked to each other but in no way from the other features. This result is correct since the date and epoch are set with an arbitrary scale. Also, as expected, the Perihelion Epoch and Osculation Time are linked, but there is no dependence on the orbit parameters. Furthermore, the mean motion is conditionally independent with relative velocity, but correctly, this is dependent on perihelion distance.

@@

\section{Machine learning algorithms}

The performances of mgm model will be now compared with the obtained from other ML algorithms. These include the random-forest (as implemented in the RandomForest package \cite{rfor}), the support vector machines (as implemented in the e1071 package \cite{dimitriadou2008misc}) , the quadratic discriminant analysis (as implemented in the MASS package \cite{MASS} and the logistic regression (as implemented in stats package \cite{stats}). Their performances are reported in Fig. \ref{CF_ML} and \ref{ROC_mgm} as well in the Tab. \ref{phi_values}. This comparison shows that the RF and the SVM outperform the mgm graphical method while the logistic one has similar performances. Thus, one can ask the advantage of using a graphical method instead of a random forest or an SVM since its performances seem lower. The answer is the interpretability of the model provided: the random forest, at least, can provide variable rank importance as reported in Fig. \ref{RF_Importance} (note that apart from the Min Orbit intersection, there is a slight reshuffling in the feature importance to the one established by the MI in fig \ref{Mutual_information}; however apart from this is a black-box as the SVM, the logistic regression and the QDA. The probabilistic graph, providing conditional independences, gives the user an interpretable model whose properties can also be compared, discussed, and validated with the theory. Thus the model developed in this way allows a more scientific evaluation of the black-box ones. In the author's view, this characteristic compensates for the lack of predictive power regarding the RF or the SVM methods.

@@

  \begin{figure}[ht]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/ROC_RF.pdf}
    \subcaption{Random Forest}
    \vspace{4ex}
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/ROC_SVM.pdf}
    \subcaption{SVM}
    \vspace{4ex}
  \end{minipage} \\
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/ROC_QDA.pdf}
    \subcaption{QDA}
    \vspace{4ex}
  \end{minipage}%%
    \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/ROC_logistic.pdf}
    \subcaption{Logistic}
    \vspace{4ex}
  \end{minipage}%%
\caption{ROC curves for a selected set of ML algorithms. Plot obtained with caret and ggplot2 package \cite{kuhn2008building},ggplot2}}
\label{ROC_ML}
  \end{figure}

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{Figures/RF_Importance.pdf}
\caption{Variable importance according to the random forest algorithm as implemented in \cite{rfor} package. Plot obtained with ggplot2 package \cite{ggplot2}}
\label{RF_Importance}
\end{center}
\end{figure}

>>>>

\section*{Abstract} The machine learning algorithms provide a promising approach to classify and predict natural or social phenomena. Differently to the theoretical approach, in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., a symmetry, the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynmanlectures}), the machine learning algorithms start from very general assumptions/expressions and then their parameters, are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}. The first approach allows obtaining interpretable predictions, the second one good forecast with a reduced effort. However, the toll that has to be paid in this case is the low interpretability \cite{russell2010artificial,murphy2012machine}. Since they provide the list of conditional independences/dependencies, the graphical methods represent a good compromise between the need for interpretability and the forecasts obtained without developing a general theory. In order to prove the validity of this statement, I considered a dataset for which the theory that interconnects its feature was known: in particular, I chose the asteroids hazardousness as provided by CNEOS \cite{cneos+nasa} and published on Kaggle \cite{kaggle_dataset}. The outcomes of the probabilistic methods for predicting the asteroid hazardousness were compared with the ones provided by the theory and with the ones of Random Forest (RF), Support Vector Machines (SVM), and Logistic Regression and Quadric Discriminant analysis (QDA). The results show that the forecast performances of probabilistic methods are better than QDA and almost equal to the logistic regression but lower to RF and SVM. However, since the list of conditional dependencies correctly reflects the laws of celestial mechanics \cite{murray1999solar}, it can be said that in this case, the probabilistic methods provide an interpretable and correct explanation of their mechanism. Therefore, contrary to other machine learning algorithms, such methods can be fully validated scientifically.

@@

\chapter{Introduction} The description and the forecast of physical \footnote{Where here physical should be intended as mesuarable} phenomena can be attacked with two different approaches: starting from a restricted set of principles/axioms, one can formulate theoretical models that provide equations that describe such phenomena. Such approach will be called here \textit{ab-initio}. On the other side, one can start from data and fit them into some general models (with fixed or not fixed number of parameters). These methods will be called here the machine learning (ML) methods. In the ab-initio case, one can fully explain the laws obtained and an overall idea of why nature works in this way. The toll that has to be paid in this case is that, since the phenomena are much more complicated than the starting principles, a consistent part of them may be excluded by the assumption made at the beginning. Furthermore, the calculation of the solution given the dynamical equation can be computationally expensive\footnote{This is the case, for instance, of the Hartree-Fock method or Density Functional Theory: such methods rewrite, with some approximations, the Schroedinger equation into a much more computationally affordable way. However, also, in this case, the solution for solids can be costly \cite{martin_2004}}. On the other side, the ML methods, since they make much more general assumptions and work directly on data, are not limited to a particular class of phenomena. Therefore they do not require the development of a specific theory as the \textit{ab-initio} one: the same model can be applied to exoplanets habitability as well to credit risk evaluation. The drawback of this approach is that the underlying mechanism by which a forecast is preferred to another one may not be interpretable. Such dichotomy between these two methods can be explained with an example taken from astronomy: at the beginning of XX century, astronomers knew that Mercury follows an elliptical orbit that, instead of being fixed, rotates around the Sun as shown in Fig. \ref{Perihelion_precession2} Such phenomenon is called \textit{Precession of the perihelion of Mercury}. The crucial point is that there was no way to explain from the Newton gravity theory such phenomenon. At this point, scholars had two paths: describe its motion with an empirical law (perhaps with an empirical modification of Newton's gravity law) or reformulate the Newton gravity from scratch. The second approach was the one followed by A.Einstein with his general relativity theory (further details are given in the A.Zee textbook \cite{zee2013einstein}). The graphical methods represent a good compromise between these two approaches: they can be used for every physical phenomenon, and, since they provide conditional dependencies/independences, they also give an elegant and intuitive representation of their mechanisms. Therefore, if the theory of the involved process is known, one can use this class of algorithms and compare its finding with the theory. In this way, one can evaluate the quality of the forecasts and the model provided by the graphical methods. The present work aims to perform such analysis for a dataset that contain the features and the hazardousness \footnote{For Earth} of asteroids. For this dataset, indeed, the relationships between the various features are known. The present work is organized as follows: first, a brief introduction about the basic theoretical concepts of the statistical methods here used will be provided, then the dataset will be described, and finally, the principal results here obtained with graphical methods will be presented as compared with other types of machine learning algorithms. Furthermore, a recap of the celestial mechanics will be provided in Appendix A. This theory explains the connections and the meaning of the different features contained in the dataset. Finally, the code used to produce the results here reported if provided in Appendix B.

@@

In this section, we are going to review the theoretical concepts
that underlie the probabilistic methods here used: these will be
exposed following the approaches of Murphy \cite{murphy2012machine} , Koller
et al. \citep{koller2009probabilistic}, Højsgaardand et al. \citep{hojsgaard2012graphical} and Russel et al. \citep{russell2010artificial}. Furthermore, a rapid overview of the main concepts of information theory will also be provided following the Cover \citep{cover2006elements}, and MacKay \citep{mackay2003information} approaches. This is because some of its concepts in the preliminary analysis of the dataset were used. Finally, we concluded this chapter with a rapid overview of algorithm interpretability, which, together with the accuracy and the
confusion matrix, is necessary for the assessment of a ML performances. On the other side, the concepts related to the celestial
mechanics here used will be described, following the Murray
approach \cite{murray1999solar} into the Appendix A.

@@

\section{Information theory}
Given an ensemble of random variables, we can quantify the amount of information that one variable contains: such quantity is called mutual information and is a crucial concept within information theory. This approach, which was implemented by Claude Shannon decades before the probabilistic modelling, represents a complementary way by which one can attack the problem of conditional dependence between random variables. Furthermore, as shown in Fig. \ref{Information_theory_connections}, this theory provides a formidable contribution to different scientific fields. Here we would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allow defining the concept of mutual information properly. The founding concept of information theory is entropy. This quantity expresses the uncertainty of a random variable. Given a random variable $X$ with alphabet (the accessible states) $\cite{cover2006elements}$ and probability mass function $p(x)=Pr\left\lbrace X=x \rbrace\; x \in \chi$ we define the entropy of $X$ as $H(X)=-\sum_{x \in X} p(x)log p(x)$ where the logarithm has to be considered with basis 2. In analogous way the joint entropy of two random variables $(X,Y)$ with a joint distribution p(x,y) is defined as \cite{cover2006elements}:
\begin{equation}
H(X,Y)=-\sum_{x\in \chi}\sum_{y\in \mathcal{Y}}p(x,y)\log p(x,y)
\end{equation}
Furthermore, we can define also the conditional entropy as \cite{cover2006elements}:
\begin{equation}
\begin{split}
H(X|Y)=& \sum_{x\in \mathcal{X} }p(x)H(Y|X=x)\\
=& -\sum_{x\in \mathcal{X}}p(x)\sum_{y\in \mathcal{Y}}p(x,y)\log p(y|x) \\
=& -E\log p(Y|X)
\end{split}
\end{equation}
The joint entropy and the conditional entropy are related by the chain rule \cite{cover2006elements}:
\begin{equation}
H(X,Y)=H(X)+H(Y|X)
\end{equation}
Such rule can be extended to to the following from \cite{cover2006elements}:
\begin{equation}
H(X,Y|Z)=H(X|Z)+H(Y|X,Z)
\end{equation}
Given a distribution q and another distribution p, one can quantify how inefficiently the second one describes the first one using the concept of relative entropy or Kullback-Leibler distance \cite{cover2006elements,mackay2003information}:
\begin{equation}
D(p||q)=\sum p(x)\log\frac{p(x)}{q(x)}
\end{equation}
As stated by the Gibbs inequality \cite{cover2006elements,mackay2003information}:
\begin{equation}
D(p||q)\geq 0
\end{equation}
this quantity can not be negative: the entropy of a random variable associated with another cannot have a degree of uncertainty lower to the quantity it aims to describe. On these bases, we are now ready to introduce the concept of mutual information. This is defined as  \cite{cover2006elements}:
\begin{equation}
\begin{split}
I(X;Y)&=\sum\sump(x,y)\log\dfrac{p(x,y)}{p(x)p(y)}=\\
&=D(p(x,y)||p(x)p(y)) \\
&=H(X)-H(X|Y)=H(Y)-H(Y|X)
\end{split}
\end{equation}
 As for the joint distribution also in this case, we have a chain rule  \cite{cover2006elements}:
\begin{equation}
I(X_{1},X_{2},...,X_{n};Y)=\sum^{n}_{i=1}I(X_{i};Y|X_{i-1},X_{i-2},...,X_{1})
\end{equation}
The concepts here reviewed, and the relations that interconnect them can be represented as shown in Fig. \ref{Entropy_MI}. Finally, we would report the data process inequality theorem that connects the information theory with the Markov chain: if we have a Markov chain, $X\rightarrow Y \rightarrow Z$  then $I(X;Y)\geq I(X;Z) $. As for the Gibbs inequality, the underlying idea is that no clever manipulation of the data can improve the inference that can be made on them \cite{cover2006elements,mackay2003information}. Otherwise, we would have a clear violation of the second principle of thermodynamics (see for instance, the Maxwell's demon \cite{feynman2018feynman}).

\section{Algorithm interpretability}

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{Figures/Information_theory_connections.jpg}
\caption{The connections of information theory with different scientific fields. Image taken from \cite{cover2006elements}}
\label{Information_theory_connections}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{Figures/Entropy_MI.png}
\caption{The relations between the entropy, conditional entropy and mutual information. Image taken from \cite{cover2006elements}}
\label{Entropy_MI}
\end{center}
\end{figure}

@@

\section{Probabilistic models} The analysis with the graphical models was started by inspecting the relations between continuous variables in the dataset. Therefore, in this first step, the binary variable \textit{Hazardous} was excluded. For this purpose, among the different methods available, we considered the \textit{Graphical least absolute shrinkage and selection operator} GLASSO as implemented in the \textit{glasso} R package \cite{friedman2008sparse,glasso}. After different tests reported in Fig. \ref{GLASSO_convergence}, we considered as the final result the graph obtained with the value of $\rho$ (the one that penalizes further connections) equal to $0.3$. Such choice is motivated by the fact that the conditional dependences/independences stated by the Celestial Mechanics are correctly reproduced with this parameter. For instance, we see that the diameter of the asteroids is independent of all features related to their motion. Such a result is meaningful since the volume/mass of the asteroids is largely lower than the mass of the earth.  Thus, as explained in Appendix A, there is no way the asteroid orbit can be modified by its mass. In addition, we see that the Close approach Date and the Epoch date are linked to each other but in no way from the other features. This result is correct since the date and epoch are set with an arbitrary scale. Also, as expected, the Perihelion Epoch and Osculation Time are linked, but there is no dependence on the orbit parameters. Furthermore, the mean motion is conditionally independent with relative velocity, but correctly, this is dependent on perihelion distance.

@@

\section{Machine learning algorithms} The performances of mgm model will be now compared with the one obtained from other ML algorithms. These include the random-forest (RF, as implemented in the RandomForest package \cite{rfor}), the support vector machines (SVM, as implemented in the e1071 package \cite{dimitriadou2008misc}), the quadratic discriminant anchiedersialysis (QDA, as implemented in the MASS package \cite{MASin reverseS} and the logistic regression (as implemented in stats package \cite{stats}). Their performances are reported in Fig. \ref{CF_ML} and \ref{ROC_mgm} as well in the Tab. \ref{phi_values}. This comparison shows that the RF and the SVM outperform the mgm graphical method while the logistic one has similar performances. Thus, one can wonder about the advantage of using a graphical method instead of a random forest or an SVM since its performances seem lower. The answer is the interpretability of the model provided: the RF, at least, can provide variable rank importance as reported in Fig. \ref{RF_Importance} (note that apart from the Min Orbit intersection, there is a slight reshuffling in the feature importance to the one estain reverseblished by the MI in fig \ref{Mutual_information}. However, apart from this intepretation, the RF is a black-box as the SVM, the logistic regression and the QDA. Conversely, the probabilistic graphs, providing conditional independences, give the user an interpretable model whose properties can also be compared, discussed, and validated with the theory. Thus the model developed in this way allows a more scientific evaluation with respect to the black-box ones. In the author's view, this characteristic compensates for the lack of predictive power regarding the RF or the SVM methods.

@@

  \begin{figure}[ht]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/ROC_RF.pdf}
    \subcaption{Random Forest}
    \vspace{4ex}
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/ROC_SVM.pdf}
    \subcaption{SVM}
    \vspace{4ex}
  \end{minipage} \\
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/ROC_QDA.pdf}
    \subcaption{QDA}
    \vspace{4ex}
  \end{minipage}%%
    \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/ROC_logistic.pdf}
    \subcaption{Logistic}
    \vspace{4ex}
  \end{minipage}%%
\caption{ROC curves for a selected set of ML algorithms. Plot obtained with the caret and the ggplot2 package \cite{kuhn2008building},ggplot2}}
\label{ROC_ML}
  \end{figure}

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{Figures/RF_Importance.pdf}
\caption{Variable importance according to the random forest algorithm as implemented in \cite{rfor} package. Plot obtained with the ggplot2 package \cite{ggplot2}}
\label{RF_Importance}
\end{center}
\end{figure}
